An Experimental Analysis of the Influence of the Precision of Parameters in an Artificial Neural Network

### 2025-04-08 presentations/02/main.tex
presentation 02

\section{Introduction}

% each frame is a 'slide'
\begin{frame}{Primer on Neural Networks}
  \begin{figure}
    \includegraphics[width=0.7\textwidth]{figures/mlp.png}
    \caption{Example png.}
  \end{figure}
\end{frame}

\begin{frame}{Single neuron view}
  \begin{figure}
    \includegraphics[width=0.96\textwidth]{figures/single-neuron.pdf}
    \caption{Example pdf.}
  \end{figure}

  You can still add more here.
\end{frame}

\begin{frame}{Gradients}
  \begin{itemize}
    \item Want to optimize the parameters $\omega$ and $b$.
    \item Need a target $y$ and error metric (loss $L$) to judge the networks predictions $\hat y$.
    \item Convex optimization is used to adjust the parameters.
    \item Many intermediate results need to be stored for efficient gradient computation.
  \end{itemize}

  \begin{align*}
    L(f(x_1, x_2)) &= (\sigma(\omega^T x + b) - y) ^2\\
                   &= (\sigma(a) - y) ^ 2\\
                   &= (\hat y - y) ^ 2
  \end{align*}
  \begin{alignat*}{3}
    \pdv{L}{\omega_1} &= \pdv{L}{\sigma} &&\pdv{\sigma}{a}      &&\pdv{a}{\omega_1}\\
                      &= 2(\hat y - y)~~ &&\odv{}{a}\sigma(a)~~ &&x_1
  \end{alignat*}

\end{frame}

\begin{frame}
  Frame without title.\\

  Example citation~\cite{Russel2020}.
\end{frame}

\begin{skipframecount}

\begin{frame}[allowframebreaks]{References}
  \printbibliography[heading=none]
\end{frame}

\appendix

\end{skipframecount}

>>>>

\begin{frame}{Recap and Today}
  \textbf{What are we doing:} Change parameter precision of ANN.\\
  \vspace{0.5cm}
  \textbf{This presentation:} Look at the effects of
  \begin{enumerate}
    \item lower precision in training and inference.
    \item lower precision only in inference.
  \end{enumerate}
\end{frame}

\begin{frame}{Recap and Today}
  \begin{itemize}
    \item Want to optimize the parameters $\omega$ and $b$.
    \item Need a target $y$ and error metric (loss $L$) to judge the networks predictions
          $\hat y$.
    \item Many intermediate results need to be stored for efficient gradient computation.
  \end{itemize}

  \begin{align*}
    L(f(x_1, x_2)) & = (\sigma(\omega^T x + b) - y) ^2 \\
                   & = (\sigma(a) - y) ^ 2             \\
                   & = (\hat y - y) ^ 2
  \end{align*}
  \begin{alignat*}{3}
    \pdv{L}{\omega_1} & = \pdv{L}{\sigma} & \cdot & \pdv{\sigma}{a}      & \cdot & \pdv{a}{\omega_1} \\
                      & = 2(\hat y - y)~~ & \cdot & \odv{}{a}\sigma(a)~~ & \cdot & x_1
  \end{alignat*}
\end{frame}

\begin{frame}{Data and Network Structure}
  \textbf{MNIST:} $28 \times 28$ pixel images of handwritten digits.

  \begin{figure}
    \includegraphics[width=0.7\textwidth]{figures/mnist.png}
    \caption{MNIST dataset example images~\cite{mnistImage}.}
  \end{figure}
\end{frame}

\begin{frame}{Data and Network Structure}
  Model evaluation is done with with 5-fold CV.\\

  \begin{figure}
    \includegraphics[width=0.8\textwidth]{figures/mlp.png}
    \caption{Structure of simple NN to classify MNIST numbers.}
  \end{figure}
\end{frame}

\begin{frame}{Results - Training and Inference}
  \begin{figure}
    \includegraphics[width=0.7\textwidth]{figures/trainingAndInference.png}
    \caption{Model accuracies - trained and evaluated at each precision.}
  \end{figure}
\end{frame}

\begin{frame}{Results - Training and Inference Times}
  \begin{figure}
    \includegraphics[width=0.7\textwidth]{figures/times.png}
    \caption{Training and evaluation times.}
  \end{figure}
\end{frame}

\begin{frame}{Results - Inference-Only Testing}
  \begin{figure}
    \includegraphics[width=0.7\textwidth]{figures/inferenceOnly.png}
    \caption{Model accuracies - trained with float64 and post-training precision reduction.}
  \end{figure}
\end{frame}

\begin{frame}{Conclusion and Outlook}
  \textbf{Interpretation:} Simple task and simple network $\rightarrow$ almost no effets.\\
  \vspace{0.5cm}
  \textbf{Next Steps:}
  \begin{enumerate}
    \item Reduce precision even further $\rightarrow$ quantize lower than float16 (half
          precision).
    \item Mixed Precision Training $\rightarrow$ different bit-widths for different parts
          of the training process.
  \end{enumerate}
\end{frame}

%%%%%%%%%%%%%%%%%%%%

\begin{skipframecount}
  \begin{frame}[allowframebreaks]{References}
    \printbibliography[heading=none]
  \end{frame}

  \appendix
\end{skipframecount}
